---
title: "Drug masking report"
author: "Haley Hunter-Zinck"
date: "`r Sys.setenv(TZ='America/Los_Angeles'); format(Sys.time(), '%B %d, %Y at %I:%M %p %Z')`"
output: html_document
params:
  cohort: x
  site: x
  date: x
  auth: x
---

## Drug masking report

This report checks for potentially unmasked investigational drugs with the following checks:

(@) Investigational drugs should have a duration equal to 1
(@) For investigational drugs, the other drug name field should be empty
(@) Investigational drugs should be part of a clinical trial
(@) Any unmasked drug should have a HemOnc FDA approval date before the supplied index date (date of latest curation)
(@) Redacted patient records, which may require additional scrutiny

```{r setup, include=FALSE}
library(synapser)
library(glue)
library(yaml)
library(knitr)
library(dplyr)

config <- read_yaml("config.yaml")

knitr::opts_chunk$set(echo = F)

```

```{r functions, include = F}

#' Get the name of a Synapse entity. 
#' 
#' @param synapse_id Synapse ID string
#' @return String representing entity name
#' @example get_synapse_entity_name("syn12345")
get_synapse_entity_name <- function(synapse_id) {
  return(synGet(synapse_id, downloadFile = F)$properties$name)
}

get_synapse_folder_children <- function(synapse_id) {
  
  
  ent <- as.list(synGetChildren(synapse_id))
  
  children <- c()
  for (i in 1:length(ent)) {
    children[ent[[i]]$name] <- ent[[i]]$id
  }
  
  return(children)
}

get_synid_report <- function(cohort, site, date, synid_folder_qc, map_cohort = map_cohort) {
  
  # get cohort subfolder synapse id
  synid_children_cohort <- get_synapse_folder_children(synid_folder_qc)
  synid_folder_cohort <- as.character(synid_children_cohort[map_cohort[[cohort]]])
  
  # get synapse ID of folder labeled "qa_drug_review_qa_manager_signoff"
  synid_children_review <- get_synapse_folder_children(synid_folder_cohort)
  synid_folder_review <- as.character(synid_children_review["qa_drug_review"])
  
  # get synapse ID of folder for most recent previous review
  synid_children_date <- get_synapse_folder_children(synid_folder_review)
  synid_folder_date <- as.character(synid_children_date[glue("{date} Reports")])
  
  synid_children_report <- get_synapse_folder_children(synid_folder_date)
  synid_file_report <- as.character(synid_children_report[glue("unmasked_drugs_review_{site}_{cohort}_{date}.csv")])
  
  return(synid_file_report)
}

get_latest_curation_date <- function(synapse_id, cohort) {
  query <- glue("SELECT MAX(curation_dt) FROM {synapse_id} WHERE cohort = '{cohort}'")
  dt <- as.character(as.data.frame(synTableQuery(query, includeRowIdAndRowVersion = F))[1])
  return(dt)
}

#' Extract personal access token from .synapseConfig
#' located at a custom path. 
#' 
#' @param path Path to .synapseConfig
#' @return personal acccess token
get_auth_token <- function(path) {
  
  lines <- scan(path, what = "character", sep = "\t", quiet = T)
  line <- grep(pattern = "^authtoken = ", x = lines, value = T)
  
  token <- strsplit(line, split = ' ')[[1]][3]
  return(token)
}

#' Override of synapser::synLogin() function to accept 
#' custom path to .synapseConfig file or personal authentication
#' token.  If no arguments are supplied, performs standard synLogin().
#' 
#' @param auth full path to .synapseConfig file or authentication token
#' @param silent verbosity control on login
#' @return TRUE for successful login; F otherwise
#' Override of synapser::synLogin() function to accept 
#' custom path to .synapseConfig file or personal authentication
#' token.  If no arguments are supplied, performs standard synLogin().
#' 
#' @param auth full path to .synapseConfig file or authentication token
#' @param silent verbosity control on login
#' @return TRUE for successful login; F otherwise
synLogin <- function(auth = NA, silent = T) {
  
  # default synLogin behavior
  if (is.na(auth)) {
    syn <- synapser::synLogin(silent = silent)
    return(T)
  }
  
  token = auth
  
  # extract token from .synapseConfig
  if (grepl(x = auth, pattern = "\\.synapseConfig$")) {
    token = get_auth_token(auth)
    
    if (is.na(token)) {
      return(F)
    }
  }
  
  # login
  syn <- tryCatch({
    synapser::synLogin(authToken = token, silent = silent)
  }, error = function(cond) {
    return(F)
  })
  
  if (is.null(syn)) {
    return(T)
  }
  return(syn)
}

```

```{r parameters, include = F}

cohort <- params$cohort
site <- params$site
date <- params$date   
auth <- params$auth

msg_pass <- "PASS: no such records found"
msg_fail <- "WARNING: issue found with {n_record} record(s)"

login_status <- synLogin(auth = auth)

synapse_id <- get_synid_report(cohort = cohort, 
                                    site = site, 
                                    date = date, 
                                    synid_folder_qc = config$synapse$qc$id,
                                    map_cohort = config$map$cohort_name)
date_curation <- get_latest_curation_date(synapse_id = config$synapse$curation$id, 
                                          cohort = cohort)
```

```{r print params, echo = F}
print(glue("Cohort: {cohort}"))
print(glue("Site: {site}"))
print(glue("Date of upload: {date}"))
print(glue("Index date (date of latest curation): {date_curation}"))
print(glue("Report data: {get_synapse_entity_name(synapse_id)} ({synapse_id})"))
```

```{r read, include = F}
data <- read.csv(synGet(synapse_id)$path, stringsAsFactors = F)
```


## Investigational drug duration

The following records have administered drugs which are marked as investigational but a duration that is not equal to 1 day:  

```{r duration}
idx_duration <- which(data[,"investigational_duration"] != "ok")
if(length(idx_duration)) {
  n_record <- length(idx_duration)
  print(glue(msg_fail))
  capt <- glue("Records with investigational drugs with duration not equal to 1 day (n={n_record})")
  kable(data[idx_duration,c("cohort", "record_id", "regimen_number")], capt = capt, row.names = F)
} else {
  print(msg_pass)
}

```

## Investigational drug with other name

The following records have administered drugs which are marked as investigational but also a drug notated in the other columns:

```{r other}
idx_other <- which(data[,"investigational_other"] != "ok")
if(length(idx_other)) {
  n_record <- length(idx_other)
  print(glue(msg_fail))
  capt <- glue("Records with investigational drugs but other drug noted (n={n_record})")
  kable(data[idx_other,c("cohort", "record_id", "regimen_number")], capt = capt, row.names = F)
} else {
  print(msg_pass)
}
```

## Investigational drug not in clinical trial

The following records have administered drugs which are marked as investigational but not noted as being part of a clinical trial:

```{r trials}
idx_ct <- which(grepl(data[,"drugs_in_regimen"], pattern = "Investigational Drug") & data[,"drugs_ct_yn" ] == "No")
if(length(idx_ct)) {
  n_record <- length(idx_ct)
  print(glue(msg_fail))
  capt <- glue("Records with investigational drugs but not in clinical trial (n={n_record})")
  kable(data[idx_ct,c("cohort", "record_id", "regimen_number", "drugs_in_regimen", "drugs_ct_yn")], capt = capt, row.names = F)
} else {
  print(msg_pass)
}
```

## Unapproved or unknown FDA status for drugs in a regimen

The following records have administered drugs which do not have a valid FDA approval date according to the HemOnc ontology and supplied index date:

```{r fda}
idx_fda <- which(is.element(data[,"fda_status"], c("unapproved", "unknown")))
if(length(idx_fda)) {
  
  # sort by fda status for easier reading
  to_kable <- data[idx_fda,c("cohort", "record_id", "regimen_number", "drugs_in_regimen", "drugs_ct_yn", "fda_status")]
  to_kable <- to_kable %>% 
    arrange(fda_status)
  
  n_record <- length(idx_fda)
  print(glue(msg_fail))
  capt <- glue("Records have no FDA approval date (n={n_record})")
  kable(to_kable, capt = capt, row.names = F)
} else {
  print(msg_pass)
}
```

## Regimens for patients with redacted data

The following records are associated with redacted patients (age above 89 years) and may require additional scrutiny, especially with regard to the drug duration:

```{r redacted}

idx_redac <- which(data[,"age_greater_89"] == 1)

if(length(idx_redac)) {
  
  # sort by fda status for easier reading
  to_kable <- data[idx_redac,c("cohort", "record_id", "regimen_number", "drugs_in_regimen", "drugs_ct_yn", "fda_status")]
  to_kable <- to_kable %>% 
    arrange(fda_status)
  
  n_record <- length(idx_redac)
  print(glue(msg_fail))
  capt <- glue("Records are redacted (n={n_record})")
  kable(to_kable, capt = capt, row.names = F)
} else {
  print(msg_pass)
}
```
