---
title: "Summary of case selection"
author: "Haley Hunter-Zinck, Xindi Guo"
date: "`r Sys.setenv(TZ='America/Los_Angeles'); format(Sys.time(), '%B %d, %Y at %I:%M %p %Z')`"
output: html_document
params:
  phase: x
  cohort: x
  site: x
---

```{r setup, include=FALSE}

library(synapser)
synLogin()
library(yaml)
library(glue)
library(dplyr)
library(RColorBrewer)
library(RCurl)
library(jsonlite)

# rmarkdown knit setup
knitr::opts_chunk$set(echo = F)

# read in all global parameters
config <- read_yaml("config.yaml")

# user specified parameters
phase <- as.character(params$phase)
cohort <- params$cohort
site <- params$site

# file names
file_matrix <- tolower(glue("{cohort}_{site}_phase{phase}_eligibility_matrix.csv"))
file_selection <- tolower(glue("{cohort}_{site}_phase{phase}_case_selection.csv"))
```

```{r functions, include = F}

#' Traverse the oncotree hierarchy to get all descendant codes of a root code.
#' 
#' @param root Root oncotree code
#' @param url URL of holding JSON object of oncotree hierarchy
#' @return All descendant codes of root in a vector of strings.
#' @example 
#' traverse(root = "BLADDER")
traverse <- function(root, oncotree) {
  
  desc <- c()
  
  children = oncotree$code[which(oncotree$parent == root)]
  
  if(!length(children)) {
    return(root)
  }
  
  for (child in children) {
    desc <- append(desc, traverse(child, oncotree))
  }
  
  return(c(root, desc))
}
```

```{r read, include = F}

eligibility_matrix <- read.csv(file_matrix)
case_selection <- read.csv(file_selection)

```

# Parameters

**Phase**: `r phase`

**Cohort**: `r cohort`

**Site**: `r site`

# Eligibility

Eligible samples are defined with following criteria:

1. Sample has an allowed OncoTree code
    * Allowed codes: `r paste0(config$phase[[phase]]$cohort[[cohort]]$oncotree$allowed_codes, collapse = ", ")`
2. Patient contributing sample was within the allowable age range at the time of sequencing
    * Minimum age: `r config$default$global$age_min`
    * Maximum age: `r if (!is.null(config$phase[[1]]$cohort[[cohort]]$age_max)) config$phase[[1]]$cohort$BrCa$age_max else config$default$global$age_max`
3. Sample was sequenced between allowed date range
    * Minimum sequencing date: `r config$phase[[phase]]$cohort[[cohort]]$date$seq_min`
    * Maximum sequencing date: `r config$phase[[phase]]$cohort[[cohort]]$date$seq_max`
4. Patient was sequenced in the year of death or before. _(Patient with last contact int before age of sequencing needs to be reviewed by the QA manager)_
5. Patient was not already included in the cohort in a previous BPC release.
6. Patient or sample was not previously excluded for quality control or other reasons.

```{r eligibility}

query <- glue("SELECT COUNT(*) FROM {config$synapse$main_patient$id} WHERE CENTER = '{site}'")
n_main_genie <- as.integer(as.data.frame(synTableQuery(query, includeRowIdAndRowVersion = T)))

print(glue("Total number of main GENIE cases for site: {n_main_genie}"))
print(glue("Total number of BPC eligible cases: {nrow(case_selection)}"))
```

# Identify categories of eligible cases

Eligible cases fall into the following categories:

* production
    * sdv: cases that undergo source data verification (SDV).  note: all pressure cases undergo SDV and the total number of cases required for SDV includes the pressure cases
    * pressure: pressure cases that also undergo source data verification (SDV)
    * irr: cases that undergo double abstraction to calculate inter-rater reliability (IRR) 
    * review: cases that need to be reviewed by QA manager due to last contact before age of sequencing
* extra: extra eligible cases not included in the production cohort

```{r categories}
print(glue("Target number of production cases: {case_selection %>% filter(category == 'production') %>% count()}"))
print(glue("Target number of SDV cases (including pressure): {case_selection %>% filter(sdv == 'sdv') %>% count()}"))
print(glue("Target number of pressure cases: {case_selection %>% filter(pressure == 'pressure') %>% count()}"))
print(glue("Target number of IRR cases: {case_selection %>% filter(irr == 'irr') %>% count()}"))
print(glue("Extra eligible cases: {case_selection %>% filter(category == 'extra') %>% count()}"))
print(glue("Cases to be reviewed: {sum(case_selection$review)}"))
```

# Potential codes 
AACR supplies a list of eligible OncoTree codes.  Here, we traverse the branches to list other potential codes that lie beneath a relevant OncoTree root code but are not included in the list of allowable codes for cohort eligibility.  

```{r potential codes}

oncotree <- fromJSON(getURL("http://oncotree.mskcc.org/api/tumorTypes"), flatten = T)
root <- config$phase[[phase]]$cohort[[cohort]]$oncotree$root
descendants <- traverse(root = root,
                     oncotree = oncotree)

selected <- config$phase[[phase]]$cohort[[cohort]]$oncotree$allowed_codes
excluded <- setdiff(descendants, selected)

selected_str <- paste0(selected, collapse = ", ")
excluded_str <- paste0(excluded, collapse = ", ")
print(glue("Selected codes: {selected_str}"))
print(glue("Selected codes under root {root}: {paste0(intersect(selected, descendants), collapse = ', ')}"))
print(glue("Selected codes NOT under root {root}: {paste0(setdiff(selected, descendants), collapse = ', ')}"))
print(glue("Non-selected codes under root {root}: {excluded_str}"))

```
